---
title: "introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview
spaSim3D (**spa**tial **Sim**ulator 3D) in a simulator of 3D tissue spatial 
data. It includes a diverse range of functions to meet your simulation needs.
This includes generation and customisation of background cells in the 3D tissue.
This also includes generation and customisation of various cell clusters, such
as spheres, ellipsoids, cylinders and network clusters with or without cellular
rings and double rings. Collectively, these tools can help to mimic a realistic
3D biological tissue.

# Installing
spaSim3D is available for installation via our GitHub.

```{r install, eval = FALSE}
### Install from GitHub ###
# install.packages("devtools")
# devtools::install_github("TrigosTeam/spaSim3D")

library(spaSim3D)
```

# Visual example
```{r}
# spe_metadata <- spe_metadata_background_template("random")
# spe_metadata <- spe_metadata_cluster_template("regular", "sphere", spe_metadata)
# spe_metadata <- spe_metadata_cluster_template("regular", "network", spe_metadata)
# spe_metadata <- spe_metadata_cluster_template("ring", "ellipsoid", spe_metadata)
# spe_clusters <- simulate_spe_metadata3D(spe_metadata, 
#                                         plot_image = T, 
#                                         plot_cell_types = c("Tumour", "Immune", "Immune1"),
#                                         plot_colours = c("orange", "skyblue", "blue"))
#

```

