---
title: "Simulate with metadata functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulate with metadata functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

One approach that spaSim3D uses to generate simulations is through creating and
updating simulation metadata. It works by first obtaining all the desired
parameters for a simulation in a separate metadata output, and then feeding this 
into  another function to generate the 3D simulation.

# Templates for spaSim3D metadata
First, you need to generate templates of the metadata. These templates come in 
the form of a list of all the different parameters needed. You can update these 
parameters after.

You should first start off with background metadata, which contains all 
parameters needed for the background cells of the 3D tissue. The background type
can either be set to "random" or "ordered". These types emulate the different
background cell arrangements seen in 3D tissues and are explained in detail in 
the simulate_with_classic_functions.Rmd vignette.

Then, you can add cluster metadata, which would contain parameters for the cell
clusters you want to add to your 3D tissue. Repeat this multiple times for
multiple clusters. It requires the user to specify the cluster type, which can
be set to "regular", "ring" or "double ring", and the cluster shape, which can
be set to "sphere", "ellipsoid", "cylinder" or "network". These different 
cluster types and cluster shapes are able to capture the structures seen in 
real 3D tissues and are explained in detail in the 
simulate_with_classic_functions.Rmd vignette.

```{r}
library(spaSim3D)

# Get background metadata
background_metadata <- spe_metadata_background_template(background_type = "random",
                                                        original_spe_metadata = NULL)
 
# Add cluster metadata (using background metadata as background)
# This will be cluster_1 in the metadata
cluster_metadata <- spe_metadata_cluster_template(cluster_type = "regular", 
                                                  shape = "sphere", 
                                                  original_spe_metadata = background_metadata)

# This will be cluster_2 in the metadata
cluster_metadata <- spe_metadata_cluster_template(cluster_type = "ring", 
                                                  shape = "ellipsoid", 
                                                  original_spe_metadata = cluster_metadata)

# This will be cluster_3 in the metadata
cluster_metadata <- spe_metadata_cluster_template(cluster_type = "double ring", 
                                                  shape = "cylinder", 
                                                  original_spe_metadata= cluster_metadata)

```

# Updating the templates
The templates contain parameters equal to arbitrary values. However, it is up to 
the user to change the parameters to their specifications. For example, changing
the 3D tissue size, or increasing the size of the clusters, or changing the
centre position of the clusters, etc.

There are some parameters that should never update. For the background, don't
update the background_type - if you want to switch from random to ordered or 
vice versa, remake the template. For the clusters, don't update the cluster_type
or shape, remake the template instead. It's also important that you don't make
new parameters, and only update the existing parameters found in the templates.

For description about the parameters that you can update, this is lazy answer,
but please see the simulate_with_classic_functions.Rmd vignette and do
help(whatever_simulation_function_in_the_vignette) to get a breakdown of what 
each parameter means.

This part is a bit tedious, however, for a large scale simulation project,
often these parameters are 'set and forget' - figure what parameters you want to
work with at the start, and you won't really need to update it later.
```{r}
# Change background metadata
cluster_metadata$background$n_cells <- 30000
cluster_metadata$background$width <- 450
cluster_metadata$background$minimum_distance_between_cells <- 2
cluster_metadata$background$cell_types <- c("Others", "Immune")
cluster_metadata$background$cell_proportions <- c(0.99, 0.01)

# Change cluster metadata
cluster_metadata$cluster_1$radius <- 80
cluster_metadata$cluster_1$cluster_cell_proportions <- c(0.9, 0.1, 0)

cluster_metadata$cluster_2$ring_width <- 10 

cluster_metadata$cluster_3$outer_ring_width <- 5


```


# Simulating with the templates
After you've updated the parameters to your liking, simulate :)
```{r}
# Simulating
clusters <- simulate_spe_metadata3D(cluster_metadata, plot_image = FALSE)

# Plot
plot_cells3D(clusters,
             plot_cell_types = c("Others", "Tumour", "Immune", "Immune1", "Endothelial"),
             plot_colours = c("lightgray", "orange", "skyblue", "lightgreen", "tomato"))
```



