---
title: "simulate with interactive functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{simulate with interactive functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

spaSim3D has two interactive functions which allow the user to generate 3D
tissue simulations at their own pace. When running each function, you will be
asked to input specific inputs directly into the console to simulate 3D tissues.
The meaning of each input should be clearly described as you work through the 
functions, so I won't explain them in this vignette.

Start by generating the 'background' of cells for your 3D tissue using 
simulate_background_interactive().

Then, generate 3D cell clusters on top of your background using
simulate_cluster_interactive().

That's it. You can keep adding cell clusters, or start from scratch.

I think these functions are great for users who have only just encountered this
package for the first time. However, for large scale simulation projects, an
interactive approach doesn't seem logical, and I would recommend the other 
vignettes to see other approaches instead.

```{r}
### First, generate the background
# background_spe <- simulate_background_interactive()

### Add a cluster to the background
# cluster_spe <- simulate_cluster_interactive(background_spe)

### Add another cluster
# cluster_spe <- simulate_cluster_interactive(cluster_spe)
```

