---
title: "simulate with classic functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{simulate with classic functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

spaSim3D is an 3D adaptation of the original spaSim. Hence, these 'classic
functions' are the functions which are literally the 3D equivalent of the
simulation functions used in the original spaSim.

Do I recommend using these functions - No. I think inputs can get quite messy
when using these functions. I would recommend using the metadata functions
instead, which I think are more cleaner, but, you do you.

Anyway, let's go through each one :)


# Simulate background cells
spaSim3D had two methods to simulate background cells. Both methods involve
creating a 3D window and filling it with 'background cells' - blank cells with
no cell type identity. The methods differ in how the cells are placed in the
background:

1. simulate_random_background_cells3D results in a 3D tissue where the cells are
positioned randomly, following a Poisson distribution. This is designed to mimic
3D tumour tissues where the cells often exhibit a random arrangement.

2. simulate_ordered_background_cells3D results in a 3D tissue where the cell are
positioned in an organised grid pattern such that adjacent cells are equidistant
from each other. This is designed to mimic healthy 3D tissues where the cells
often exhibit a uniform arrangement.

```{r}
library(spaSim3D)

# 1. Simulate random background of cells
random_background <- simulate_random_background_cells3D(n_cells = 10000,
                                                        length = 100,
                                                        width = 100,
                                                        height = 100,
                                                        minimum_distance_between_cells = 0.5,
                                                        background_cell_type = "Others",
                                                        plot_image = FALSE)

# Plot - apparently the plot_image parameter doesn't like to show the plot when knitting to html
plot_cells3D(random_background, 
             plot_cell_types = "Others", 
             plot_colours = "lightgray")


# 2. Simulate ordered background of cells
ordered_background <- simulate_ordered_background_cells3D(n_cells = 10000,
                                                          length = 100,
                                                          width = 100,
                                                          height = 100,
                                                          jitter_proportion = 0,
                                                          background_cell_type = "Others",
                                                          plot_image = FALSE)

# Plot
plot_cells3D(ordered_background, 
             plot_cell_types = "Others", 
             plot_colours = "lightgray")

```

# Simulate mixing of the background cells
The background functions simulate a background of a blank, undefined cell type.
If users which to assign cell types randomly to the background cells, they can
use the mixing function. 

The user inputs a selection of cell types and their desired proportions. Then, 
each individual cells is reassigned to these chosen cell types with a 
probability that directly corresponds to the desired proportion of that cell 
type.

```{r}
## Simulate mixing

# The background will have 50% Other, 25% Immune and 25% Tumour cells.
mixed_background <- simulate_mixing3D(random_background,
                                      cell_types = c("Others", "Immune", "Tumour"),
                                      cell_proportions = c(0.5, 0.25, 0.25),
                                      plot_image = FALSE)

# Plot
plot_cells3D(mixed_background,  
             plot_cell_types = c("Others", "Immune", "Tumour"),
             plot_colours = c("lightgray", "skyblue", "orange"))

```


# Simulate clusters
Once you've generated the background of cells, you want to simulate clusters
within the background. These clusters can mimic clusters of tumour cells seen
in cancer tissues, or blood vessels, or immune networks. If you are creative
enough, you can use these clusters to simulate animal embryos.

spaSim3D allows you to simulate 4 different shapes of clusters: spheres,
ellipsoids, cylinders and networks. Each has their own set of parameters. E.g.
A sphere has radius and a centre coordinate. You can use 
help(whatever_simulation_function) to get a breakdown of what each parameter 
means.

Hence, the 4 different simulation functions to use to generate these 4 different
shaped cell clusters + 1 function that combines these 4 functions.

As stated before, the parameters for each shape will differ, but they all share
common parameters: cluster_cell_types and cluster_cell_proportions. These are
similar to the cell_types and cell_proportions used for the background
simulation functions - specify what cell types and the corresponding proportions 
you want for each cluster, and the cells within the cluster will be reassigned
to these cell types with a probability equal to the desired proportion of each
cell type.


```{r}
## Simulate clusters

# Use the simulate_clusters3D to simulate numerous and different types of clusters at once.
# This will simulate a sphere, cylinder and an ellipsoid.
clusters <- simulate_clusters3D(mixed_background,
                                cluster_properties = list(
                                    C1 = list(
                                        shape = "sphere",
                                        cluster_cell_types = c("Tumour", "Immune", "Others"),
                                        cluster_cell_proportions = c(0.55, 0.4, 0.05),
                                        radius = 25,
                                        centre_loc = c(40, 40, 40)
                                    ),
                                    C2 = list(
                                        shape = "cylinder",
                                        cluster_cell_types = c("Endothelial", "Others"),
                                        cluster_cell_proportions = c(0.95, 0.05),
                                        radius = 10,
                                        start_loc = c(0, 0, 0),
                                        end_loc   = c(20, 20 , 100)
                                    ),
                                    C3 = list(
                                        shape = "ellipsoid",
                                        cluster_cell_types = c("Tumour", "Immune", "Others"),
                                        cluster_cell_proportions = c(0.65, 0.3, 0.05),
                                        radii = c(15, 20, 25),
                                        centre_loc = c(70, 70, 70),
                                        axes_rotation = c(0, 0, 0)
                                    )
                                ),
                                plot_image = FALSE)

# Plot
plot_cells3D(clusters,  
             plot_cell_types = c("Others", "Immune", "Tumour"),
             plot_colours = c("lightgray", "skyblue", "orange"))

```


# Simulate rings

```{r}

```


# Simulate double rings

```{r}

```

